Combining information from photometric and spectroscopic data:\\ Don't waste your time in spectrophotometric calibration!

### 2014-04-30 tex/ms.tex
redshift notation change
more sampling text, start on results desc

\begin{abstract}
In a typical spectral-fitting project,
  there tend to be a small number of bands of well-calibrated photometry
  and thousands of poorly calibrated spectroscopic pixels.
Good spectroscopic calibration is both expensive and difficult,
  but naive weighted least-square fitting will weight
  the thousands of spectroscopic pixels far higher overall
  than the few photometric bands.
Here we present a flexible and general spectroscopic calibration model,
  which can be fit simultaneously along with whatever spectral parameters.
The model involves a polynommial calibration vector to capture large-scale calibration issues
  and a Gaussian Process to capture small-scale wiggles.
We show that, in this framework, the quality of some kinds of spectral parameter fits
  is not a strong function of the quality of the spectroscopic calibration;
  that is, for many scientific goals there is no scientific reason
  to obtain good---or really any---spectrophotometric calibration.
In particular, for stellar population fitting, high signal-to-noise data,
  and good photometry,
  we show that raw counts are just as good as calibrated counts for making inferences about the
  parameters of interest.
This obviation of calibration overheads will pay \emph{all yall mad cheddar},
  at least if yall employ full cost accounting.
\end{abstract}

@@

To demonstrate the effectiveness of our spectrophotometric calibration
strategy we will infer the parameters $\theta$ from the spectrum both
\emph{before} and \emph{after} it has been spectrophotometrically
calibrated using typical techniques and ancillary spectrophotometric
calibration data. ( CITE NC)

We model the stellar cluster as a single age SSP. The spectrum of a
star cluster under this assumption is given by
\begin{eqnarray}\label{eq:MeanModel} 
\mu(\theta, \lambda(1+z)) & = &
\frac{m_*}{4\pi d_L^2(1+z)} G(\lambda,\sigma_v) \otimes 
S(\lambda, t, Z [, \beta]) e^ {-k_\lambda\tau_V} 
\\ 
\theta & \equiv & [m_*, t, Z, \tau_V, \sigma_v, z, d_L, \beta]
\quad ,
\end{eqnarray} where the parameters $\theta$ include include the age
$t$, metallicity $Z$, stellar mass $m_*$, dust attenuation effective
$V$-band optical depth $\tau_V$, velocity dispersion $\sigma_v$,
luminosity distance $d_L$ and redshift $z$.  The spectrum of one solar
mass of stars that formed at time $t$ with metallicity $Z$ is given by
$S(\lambda, t, Z [, \beta])$ where the additional parameters $\beta$
may include details of the population synthesis, such as the IMF slope
and adjustments to the stellar spectra or stellar evolutionary tracks
\citep[e.g.][]{FSPS}. We will assume $\beta$ to be fixed for the
purposes of this exercise, with no loss of generality (KEEPING IMF
SLOPE CONSTANT FOR NOW).  We will also fix the shape of the
attenuation curve, $k_\lambda \equiv \tau_\lambda/\tau_V$ to the Milky
way average of \citet{CCM89} and assume that $d_L$ is known to high
accuracy from independent methods.

To model the photometry we simply project the observed frame mean
model spectrum $\mu(\theta, \lambda(1+z))$ onto the filters of
interest.  Note that we do \emph{not} include the calibration vector
$f(\alpha)$ in this projection.

BDJ WORDS ABOUT FSPS.

@@

Our sampling of the posterior is accomplished using \texttt{emcee}
\cite{dfm} after an intial round of minimization.

@@

BDJ INSERT WORDS HERE.
BDJ random issues, collected here in case we want to address them and
to keep me from cluttering up the main text: effective aperture
corrections (i.e. we're implicitly assuming photometry and
spectroscopy sample the same population).  Sky subtraction residuals.
Model resolution and instrument resolution.  Method only useful when
you have a model for the mean spectrum?  How to interpret GP
residuals, as a fn of S/N?  Photometric upper limits?

@@

All the code used in this project is available under an open-source license
  from \url{http://github.com/bd-j/something/}.
This version of the paper was generated
  from a git repository available at \url{http://github.com/bd-j/speccal/}
  with git hash \texttt{\githash\,(\gitdate)}.

>>>>

\begin{abstract}
In a typical spectral-fitting project,
  there tend to be a small number of bands of well-calibrated photometry
  and thousands of poorly calibrated spectroscopic pixels.
Good spectroscopic calibration is both expensive and difficult,
  but naive weighted least-square fitting will weight
  the thousands of spectroscopic pixels far higher overall
  than the few photometric bands.
Here we present a flexible and general spectroscopic calibration model,
  which can be fit simultaneously along with whatever spectral parameters.
The model involves a polynomial calibration vector to capture large-scale calibration issues
  and a Gaussian Process to capture small-scale wiggles.
We show that, in this framework, the quality of some kinds of spectral parameter fits
  is not a strong function of the quality of the spectroscopic calibration;
  that is, for many scientific goals there is no scientific reason
  to obtain good---or really any---spectrophotometric calibration.
In particular, for stellar population fitting, high signal-to-noise data,
  and good photometry,
  we show that raw counts are just as good as calibrated counts for making inferences about the
  parameters of interest.
This obviation of calibration overheads will pay \emph{all yall mad cheddar},
  at least if yall employ full cost accounting.
\end{abstract}

@@

To demonstrate the effectiveness of our spectrophotometric calibration
strategy we will infer the parameters $\theta$ from a spectrum both
\emph{before} and \emph{after} it has been spectrophotometrically
calibrated using typical techniques and ancillary spectrophotometric
calibration data. ( CITE NC)

We model the stellar cluster as a single age SSP. The spectrum of a
star cluster under this assumption is given by
\begin{eqnarray}\label{eq:MeanModel} 
\mu(\theta, \lambda_o) & = &
\frac{m_*}{4\pi d_L^2(1+z)} G(\lambda,\sigma_v) \otimes 
S(\lambda, t, Z [, \beta]) e^ {-k_\lambda\tau_V} 
\\ 
\lambda_o & = & (1+z)\lambda 
\\
\theta & \equiv & [m_*, t, Z, \tau_V, \sigma_v, z, d_L, \beta]
\quad ,
\end{eqnarray} where the parameters $\theta$ include the stellar mass
$m_*$, effective optical depth to dust in the $V$-band $\tau_V$,
velocity dispersion $\sigma_v$, luminosity distance $d_L$ and redshift
$z$.  The spectrum of one solar mass of stars that formed at time $t$
with metallicity $Z$ is given by $S(\lambda, t, Z [, \beta])$ where
the additional parameters $\beta$ may include details of the
population synthesis model, such as the IMF slope and adjustments to
the stellar spectra or stellar evolutionary tracks
\citep[e.g.][]{FSPS}. We will assume $\beta$ to be fixed for the
purposes of this exercise, with no loss of generality (KEEPING IMF
SLOPE CONSTANT FOR NOW).  We will also fix the shape of the
attenuation curve, $k_\lambda \equiv \tau_\lambda/\tau_V$ to the Milky
way average of \citet{CCM89} and assume that $d_L$ is known to high
precision from independent methods.

To model the photometry we simply project the observed frame mean
model spectrum (equation \ref{eq:MeanModel}) onto the filters of
interest.  Note that we do \emph{not} include the calibration vector
$f(\alpha)$ in this projection.

We calculate $\mu$ using the Flexible Stellar Population Synthesis
code \citep[FSPS][]{FSPS}.  We use the Padova isochrones
\citep{Padova} and high-resolution theoretical stellar spectra
\citep{CKC14} which allow us to probe super-solar metallicities at
early ages.  NEBULAR EMISSION - MASK OR INFER?

@@

We estimate the posterior through MCMC sampling. Our sampling of the
posterior is accomplished using \texttt{emcee} \cite{dfm} after an
intial round of minimization.  \texttt{emcee} is an implementation of
the affine-invariant ensemble sampler of \citet{GW}, and unlike
typical Metropolis-Hastings MCMC sampling does not require the
specification of step sizes in each dimension.  Our prior
distributions are in all cases top-hat functions, though more
informative priors distributions may be used.  Informative priors may
be especially appropriate for the calibration parameters $\alpha$ if
additional information about the instrument is available or for the
parameters of interest $\theta$, for example in the case where there
is knowledge of the metallicity distribution of the population from
which a given cluster is drawn.

NEEDS MASSIVE REWORDING.
In Figure \ref{fig:calibrated_spectrum} we show the spectrum resulting
from draws of the parameters $[\theta,alpha]$ from the posterior
distribution.  This figure also shows the calibrated spectrum on which
the priors were conditioned, and individual components of the model
constructed from each draw.   In Figure \ref{fig:calibrated_residuals}
we show the residuals between the data and the draws from the
posterior, for the spectra and the photometry.

In Figure \ref{fig:calibrated_triangle} we show the marginalized and
joint-pdfs for the parameters.

@@

BDJ INSERT WORDS HERE.
BDJ random issues, collected here in case we want to address them and
to keep me from cluttering up the main text: effective aperture
corrections (i.e. we're implicitly assuming photometry and
spectroscopy sample the same population).  Sky subtraction residuals.
Model resolution and instrument resolution.  Method only useful when
you have a model for the mean spectrum?  How to interpret GP
residuals, as a fn of S/N?  Photometric upper limits? Infer the
wavelength solution as well - stellar population (and telluric lines) as your arc.

@@

All the code used in this project is available under an open-source license
  from \url{http://github.com/bd-j/bsfh/}.
This version of the paper was generated
  from a git repository available at \url{http://github.com/bd-j/speccal/}
  with git hash \texttt{\githash\,(\gitdate)}.
